---
title: "Using saos API"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using saos API}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
  %\SweaveUTF8
---

[SAOS](https://saos-test.icm.edu.pl/) is a repository of judgments from Polish courts of all kinds - common courts (district, regional and appellate), administrative courts, the Supreme Court of Poland, the Constitutional Tribunal and National Appeal Chamber. The [saos](http://github.com/bartekch/saos) package is an R interface to [SAOS API](https://saos-test.icm.edu.pl/api). The API does not require an API key.

The package directly reflects the structure and functionalities of the API. There are two main parts of API: `search` and `dump`. `search` allows to search for judgments matching given query. `dump` allows direct access to whole repository. Both parts are described in detail below, preceded by an example of basic usage.

# Installation

Right now you may install the package only from [GitHub](https://github.com/bartekch/saos) with `devtools::install_github()`:
  
```{r, eval = FALSE}
install.packages("devtools", repos = "http://cran.rstudio.com/")
library(devtools)
install_github("bartekch/saos")
```

# Quick start

The core function is `search_judgments()`, which enables to search repository (through API) for judgments matching given query. Query handles a variety of parameters available in API. Following example will search for the first 10 judgments (starting from the latest) with any reference to words "dobra" and "osobiste", passed by common court:

```{r}
library(saos)
judgments <- search_judgments(all = "dobra osobiste", 
                              courtType = "COMMON",
                              sortingField = "JUDGMENT_DATE",
                              sortingDirection = "DESC",
                              limit = 10, 
                              verbose = FALSE)
class(judgments)
length(judgments)
str(judgments[[1]])
```

Search results do not include all available information about judgments. If you want to get all the details you need to use `get_judgments()` function:

```{r}
judgments_details <- get_judgments(judgments, verbose = FALSE)
```

Afterwards you could extract specific information with `extract()`:

```{r}
judges <- extract(judgments_details, "judges")
str(judges)
type <- extract(judgments_details, "judgmentType")
str(type)
date <- extract(judgments_details, "judgmentDate")
str(date)
```


# Searching

SAOS enables searching through [GUI](https://saos-test.icm.edu.pl/search) and [API](https://saos-test.icm.edu.pl/api/search). Both ways are equivalent to each other in a sense that the output from API contains exactly the same information as you would see in GUI. In particular, text of judgment is truncated. To see full record you need to take additional step, described in the next section. `search_judgments` function is a direct implementation of API search with a few enhancements. 

In general - `saos_search` always returns an object of class `saos_search`, which is basically a list of judgments. Objects of class `saos_search` could be concatenated and subset as ordinary lists.
```{r}
judgments <- search_judgments(limit = 1, verbose = FALSE)
class(judgments)
```

### Count search results

Sometimes it is useful to know the number of matched judgments without downloading everything. That is where function `count_judgments` comes in handy. It accepts exactly the same arguments as `search_judgments`, expect for meaningless in this case `limit`, `force` and `verbose`. For example to check how many judgments are in the repository right now:
```{r}
count_judgments()
```

### Details during run-time

If `verbose` argument is `TRUE` (default value) a few messages will be printed: final version of query sent to API, number of matching judgments, number of downloaded judgments, a progress bar and sometimes more. It is useful only in interactive mode really. If `verbose = FALSE` all messages are suppressed, but still a warning would be printed if generated.

### Limiting results set

An API could return only 100 judgments on one page, but you don't have to worry about this because paging mechanism is built in the package. However, number of judgments in search result could be huge, what will result in numerous requests to API, what, in turn, is time-consuming. Additionaly judgments are stored mainly as text, sometimes long, so the output could be of significant size. Therefore there are some precautionary measures implemented in the package.  

By default the search results are limited to 200. You could override it by setting `limit` argument to possitive integer or `NULL` if you want everything. But there is also second safeguard - the `force` argument indicates whether or not a hardcoded limit (200) should be ignored. By default it is set to `FALSE`, so if you are confident that you want to download everything, you have to use combination `limit = NULL, force = TRUE`. If you forget about `force` argument, maximally 200 judgments will be downloaded and warning will be generated. Consider the following  example.

Empty query, so all judgments are matched. `limit` is set to 100, so the first 100 judgments will be downloaded.
```{r}
j <- search_judgments(limit = 100, verbose = FALSE)
length(j)
```

Now we want to download more judgments, so we set `limit = 500`.
```{r}
j <- search_judgments(limit = 500, verbose = FALSE)
length(j)
```

We've got warning  - we forgot about `force` argument. We are sure we want everything, so we set `force = TRUE`. In addition we'll see how long it takes to download 500 judgments.
```{r}
system.time(j <- search_judgments(limit = 500, force = TRUE, verbose = FALSE))
length(j)
object.size(j)
```

We see that function call took around 20 seconds (may vary by a few second) and almost all of this time was spent on communication with API. The resulting set uses around 3Mb - and these are only search results, not full records.


### Query parameters

All parameters are named identically as in API, like `courtType`, `judgmentType` or `judgmentDateFrom` (whole list together with descriptions could be found in man pages for `search_judgments()`). Most of parameters are directly passed to API after simple format check. There are a few expections though.

But firstly a small utility function. You could check if the query sent to API is what you expected with `print_query` function. It does nothing but print string with final version of the query, actual part of the URL in API. This function doesn't communicate with API so it's fast. See examples:
```{r}
print_query()
print_query(judgmentDateFrom="2014-01-01")
print_query(judgeName="Maria Tyszel", judgmentDateTo="2014-06-30")
print_query(judgmentTypes = c("SENTENCE", "DECISION"))
print_query(all = "\"dobra osobiste\" -\"dobra publiczne\"")
print_query(all = list(include = "dobra osobiste",
                       exclude = "dobra publiczne"))
print_query(all = c("dobra osobiste", "kodeks cywilny"))
```


As for parameters, currently there are three, namely `all`, `legalBases` and `referencedRegulations`, that support query language partially similar to that in [Google search](https://support.google.com/websearch/answer/2466433?hl=en&ref_topic=3081620), i.e. accept following operators:

* `<space>` = implicit "AND" - when query contains multiple words separated by space then API will search for judgments that contain all these words; you shouldn't use "AND" explicitly,
* `OR` operator - API will search for judgments that contain one of the words connected by `OR`,
* `"..."` = quotation - API will search for judgments that contain whole phrase inside quotation marks; quotation has no effect if placed around single word, i.e. `"kodeks"` will give the same search results as `kodeks`; note that in `R` quotation marks need to be escaped, so query should look like `\"kodeks cywilny\"` to work properly,
* `-` sign - API will search for judgments that don't contain word preceded by `-` sign; the `-` sign has to be placed immediately before a word, without a space.

Operators could be mixed, e.g `\"dobra osobiste\" OR kodeks`, with following precedence:

1. quotation, 
2. negation,
3. `OR`, 
4. implicit AND.

`OR` takes into account only two immediate neighbours, so query `dobra osobiste OR kodeks` will search for judgments containing words "dobra" and either "osobiste" or "kodeks". However quotation has higher priority than `OR`, so query `"dobra osobiste" OR kodeks` will search for judgments with either the whole phrase "dobra osobiste" or the word "kodeks".

All above refers to parameters given as a single string, as in following examples.
```{r}
print_query(all = "dobra")
print_query(all = "dobra osobiste")
print_query(all = "dobra OR osobiste")
print_query(all = "\"dobra osobiste \"")
print_query(all = "-dobra -osobiste")
```

You could also pass parameters (those accepting query language) as a character vector or a list with any of two fields - `include` and `exclude`, which have to be character vectors or `NULL`. In case of a vector, if it is of length 2 or more, its elements will be put inside quotation marks (so one element is one phrase) and pasted together with a space (i.e. implicit AND operator).
```{r}
print_query(all = c("dobra", "osobiste"))
print_query(all = c("dobra osobiste", "kodeks cywilny", "kara"))
```

In case of a list elements of both `include` and `exclude` will be put inside quotation marks and pasted with a space, but elements of `exclude` will be additionaly preceded with `-` operator.
```{r}
print_query(list(include = c("dobra osobiste", "kodeks karny"), 
                 exclude = "kodeks cywilny"))
```

These two functionalities could be useful when programmatically sending many similar requests to API, e.g. from some set of words. Consider a following example - we've got a set of words and for each word we want to count how many judgments contain this word but don't contain any word from the rest of the set. If we want to do it manually it will be somewhat complicated and hard to read.
```{r}
words <- c("spadek", "syn", "córka", "dług")
for (i in seq_along(words)){
  query <- paste(c(words[i], paste0("-", words[-i])), collapse = " ")
  print(count_judgments(all = query))
}
```

Using advanced query parsing it is more clear what we are looking for.
```{r}
words <- c("spadek", "syn", "córka", "dług")
for (i in seq_along(words)){
  print(count_judgments(all = list(include = words[i],
                                   exclude = words[-i])))
}
```


Another special parameter is `judgmentTypes`. It accepts a character vector with any subset of a set {"DECISION", "RESOLUTION", "SENTENCE", "REGULATION", "REASONS"}. API will return judgments with type matching any from the given vector.
```{r}
print_query(judgmentTypes = "DECISION")
print_query(judgmentTypes = c("RESOLUTION", "REASONS"))
```


### Sorting results
API provides more than dozen variables, which could serve as sorting criterion. Full list could be found in man pages for `search_judgments()`. Note that these variables correspond to query parameters, but are named slightly different. By default search results are sorted by judgments ID's, which is different than in GUI, where results are sorted by date. Most useful variable for sorting is probably `JUDGMENT_DATE`. Obviously you can change direction of sorting by setting `sortingDirection` argument to `ASC` or `DESC`.  


# Getting all details about judgments

As said before, objects of class `saos_search` contain only this information, which is visible in GUI search view. It lacks some details, in particular full text of a judgment. Complete record of a judgment is accessible through API judgments view, which in turn is accessible with `get_judgments()` function in `saos` package. Function `get_judgment()` always returns an object of class `saos_judgments`, that is similar to `saos_search` but contains more data. You could use this function in two ways:

* providing search results
```{r}
j <- search_judgments(limit = 10, verbose = FALSE)
judgments <- get_judgments(j, verbose = FALSE)
class(judgments)
```
* providing ID of judgments you are interested in
```{r}
judgments <- get_judgments(1:10, verbose = FALSE)
class(judgments)
```

Setting `verbose` to `TRUE` allows displaying progress bar. 

Note that every judgments is stored in separate record, therefore function `get_judgments()` sends as many requests to API as there are judgments, what could be really time-consuming.

## Shiny progress bar

It is possible to take advantage of shiny's progress bar funcionalities. Just pass `updateProgress` function as an argument to `get_judgments`. Below there is a minimal shiny app showing how it works.
```{r, eval=FALSE}
library("shiny")
library("saos")
s <- search_judgments(limit = 50)

ui <- basicPage(
  actionButton('go', 'go'),
  textOutput("text")
)

server <- function(input, output) {
  output$text <- renderText({
    input$go
    if(input$go == 0) return()
    
    progress <- shiny::Progress$new(min = 0, max = 1)
    progress$set(message = "Downloading judgments: ", value = 0)
    on.exit(progress$close())
    
    prog <- function(value = NULL, detail = NULL) {
      progress$set(value = value, detail = paste0(100 * value, "%"))
    }
    
    invisible(get_judgments(s, verbose = FALSE, updateProgress = prog))
    return("OK")
  })
}

runApp(list(ui = ui, server = server))
```
# Dump

Another way to access complete records is to take advantage of `dump` functionality in API. It enables access to whole databases of judgments, common courts, Supreme Court chambers and enrichment tags, implemented in functions, respectively, `get_dump_judgments()`, `get_dump_courts()`, `get_dump_scChambers()` and `get_dump_enrichments()`.

### Judgments

`get_dump_judgments()` allows to download complete records of judgments. However, its functionality is rather limited. You could only choose to download judgments from the given time range and/or last modified since specified time. Function always returns an object of class `saos_judgments_dump`, which is almost identical to `saos_judgments`. `get_dump_judgments()` is much faster than `get_judgments()` because in one request it gets not one but one hundred judgments. 

Example:
```{r}
lastmonth <- get_dump_judgments(judgmentStartDate = Sys.Date() - 30,
                                judgmentEndDate = Sys.Date(),
                                verbose = FALSE)
class(lastmonth)
length(lastmonth)
```


### Courts and Supreme Court chambers

`get_dump_courts()` and `get_dump_scChambers()` download database of comon courts and Supreme Court chambers respectively. They return list of courts or chambers or, if you set `simplify` argument to `TRUE`, data frames with rows corresponding to courts or chambers. The saos package provides those two data frame as datasets `courts` and `scchambers`. 

### Enrichment tags

This is still an experimental part of the project. Enrichment tags are additional information about judgments extracted with text mining techniques, which weren't originally included in data. Therefore they are less reliable, but still could provide some interesting insight. `get_dump_enrichments()` returns a list of tags, with every element corresponding to one tag for one judgment. If you set `simplify=TRUE` the list will be coerced to data frame. Currently there are no more funcionalities available.


# Extracting

Let's says we have a list of judgments in an object of class `saos_search`, `saos_judgments` or `saos_judgments_dump`. To extract specific piece of information we may use `extract()` function. It is always used in the same way - first argument is the list of judgments and the second is the name of the field we are interested in. The names of the fields are exactly the same as in API, but case insensitive, see help pages for `extract()` for complete list. `extract()` always returns a data frame, whose first column is named `id` and contains judgments ID's. The rest of the columns depends on what field has been chosen. For example if you want to extract ID you will get a data frame with only one column. Sometimes there will be more than one row corresponding to one judgment. For example if you want to extract judges details, you will get data frame with so many rows for every judgment as there are judges involved in this judgment. See examples:

```{r}
judges <- extract(lastmonth, "judges")
str(judges)
type <- extract(lastmonth, "judgmentType")
str(type)
date <- extract(lastmonth, "judgmentDate")
str(date)
```

Note that because of ongoing development there is possibility that there are fields in API that haven't been yet implemented in the package, and vice versa. Caution is recommended.


